---
title: "TCGA PRAD"
author: "Levi Waldron"
date: "September 14, 2015"
output: html_document
---

# Load TCGA data

```{r}
if(!require(RTCGAToolbox)){
  biocLite("devtools")
  biocLite(c("limma", "RCircos", "data.table", "RCurl", "RJSONIO"))
  biocLite("Link-NY/RTCGAToolbox")
}
```

# Download TCGA PRAD data

```{r, cache=TRUE}
library(RTCGAToolbox)
rundates <- getFirehoseRunningDates()
analysisdates <- getFirehoseAnalyzeDates()
prad <- getFirehoseData("PRAD", runDate=rundates[1],
                      gistic2_Date=analysisdates[1], RNAseq_Gene=TRUE, 
        miRNASeq_Gene=TRUE, RNAseq2_Gene_Norm=TRUE, CNA_SNP=TRUE,
        CNV_SNP=TRUE, CNA_Seq=TRUE, CNA_CGH=TRUE,  Methylation=TRUE,
        Mutation=TRUE, mRNA_Array=TRUE, miRNA_Array=TRUE, RPPA=TRUE)
```

# Extract available data types


```{r}
choices <- tolower(gsub("_", "", c("RNAseq_Gene", "miRNASeq_Gene",
             "RNAseq2_Gene_Norm", "CNA_SNP", "CNV_SNP", "CNA_Seq",
             "CNA_CGH", "Methylation", "Mutation", "mRNA_Array",
             "miRNA_Array", "RPPA")))
```

```{r, cache=TRUE}
dses <- lapply(choices, function(choice) try(extract(prad, choice, 
                                            clinic=TRUE),
                                             silent=TRUE))
names(dses) <- choices
dses
```

```{r, cache=TRUE}
eset <- extract(prad, "rnaseq2genenorm")
write.csv(exprs(eset), file="prad_rnaseq.csv")
write.csv(pData(eset), file="prad_clinical.csv")
saveRDS(eset, file="prad_eset.rds")
```

To load the eset again:

```{r, cache=TRUE}
eset <- readRDS("prad_eset.rds")
```

# Make a histogram of PSA (KLK3) expression

```{r}
hist(exprs(eset["KLK3", ]))
hist(log(exprs(eset["KLK3", ])))
```

# See what clinical data are available by default:

```{r}
summary(pData(eset))
```

# Look for association between KLK3 expression and clinical PSA

```{r}
psadat <- data.frame(psa=as.numeric(as.character(eset$patient.stage_event.psa.psa_value)),
                     klk3=t(exprs(eset["KLK3", ])))
psadat.complete <- psadat[complete.cases(psadat), ]
plot(KLK3 ~ psa, data=psadat.complete, xlab="clinical PSA", ylab="KLK3 tumor expression", log="xy")
fit <- lowess(x=psadat.complete$psa, y=psadat.complete$KLK3)
lines(fit, col="red", lw=3)
```

## Is there an association between PSA in the urine and KLK3 in the tumor?

```{r}
cor.test(x=psadat$KLK3,  y=psadat$psa, method="spearman")
```


# Clinical psa as a function of ethnicity

Need to get the better race variable with complete data from Tiffany:
```{r}
table(eset$race)
boxplot(psadat$psa ~ eset$race, ylab="PSA")
```

