---
title: "TCGA OVC"
author: "Levi Waldron"
date: "March 8, 2016"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

# TCGA analysis using RTCGAToolbox

```{r, message=FALSE}
if(!require(RTCGAToolbox)){
  biocLite("devtools")
  biocLite(c("limma", "RCircos", "data.table", "RCurl", "RJSONIO"))
  biocLite("Link-NY/RTCGAToolbox")
}
library(Biobase)
library(RTCGAToolbox)
```

## Download TCGA ov data

```{r ovdownload, cache=TRUE, message=FALSE}
rundates <- getFirehoseRunningDates()
analysisdates <- getFirehoseAnalyzeDates()
ov <- getFirehoseData("OV", runDate=rundates[1],
                      gistic2_Date=analysisdates[1], RNAseq_Gene=FALSE, 
        miRNASeq_Gene=TRUE, RNAseq2_Gene_Norm=TRUE, CNA_SNP=FALSE,
        CNV_SNP=FALSE, CNA_Seq=FALSE, CNA_CGH=FALSE,  Methylation=FALSE,
        Mutation=FALSE, mRNA_Array=FALSE, miRNA_Array=TRUE, RPPA=FALSE)
```

```{r ovrnaseqextract, cache=TRUE}
eset.rnaseq <- RTCGAToolbox::extract(ov, "rnaseq2genenorm")
write.csv(exprs(eset.rnaseq), file="ov_rnaseq.csv")
write.csv(pData(eset.rnaseq), file="ov_clinical.csv")
saveRDS(eset.rnaseq, file="ov_eset.rds")
```

See what clinical data are available by default:

```{r, eval=FALSE}
summary(pData(eset.rnaseq))
```


## MicroRNA dataset
Create and save to disk:
```{r ovmirnaextract, cache=TRUE}
eset.mirna <- RTCGAToolbox::extract(ov, "miRNASeq_Gene")
saveRDS(eset.mirna, file="ov_mirna_eset.rds")
```

Don't see miR-1207, only see a few counts of 939, plenty of 637:

```{r}
mat <- exprs(eset.mirna)[c("hsa-mir-637", "hsa-mir-939", "hsa-mir-1207"), ]
library(pheatmap)
pheatmap(log2(mat + 1))
```

## plot PAX8 against hsa-mir-939

```{r}
pax8 <- exprs(eset.rnaseq)["PAX8", ]
mir939 <- exprs(eset.mirna)["hsa-mir-939", ]
pax8 <- pax8[names(pax8) %in% names(mir939)]
mir939 <- mir939[match(names(pax8), names(mir939))]
identical(names(pax8), names(mir939))
cor.test(pax8, mir939, method="spearman")  #no correlation
plot(log2(pax8+1) ~ log2(mir939+1))
```

# curatedOvarianData analysis

Load the datasets:

```{r, message=FALSE}
library(curatedOvarianData)
data(package="curatedOvarianData")
data(TCGA.RNASeqV2_eset)
data(TCGA_eset)
data(TCGA.mirna.8x15kv2_eset)
```

## PAX8 expression in TCGA RNA-seq and microarray datasets

```{r}
hist(exprs(TCGA.RNASeqV2_eset["PAX8", ]))
hist(exprs(TCGA_eset["PAX8", ]))
```

```{r}
pax8.array <- exprs(TCGA_eset)["PAX8", ]
pax8.rnaseq <- exprs(TCGA.RNASeqV2_eset)["PAX8", ]
pax8.array <- pax8.array[names(pax8.array) %in% names(pax8.rnaseq)]
pax8.rnaseq <- pax8.rnaseq[match(names(pax8.array), names(pax8.rnaseq))]
plot(pax8.rnaseq, pax8.array, xlab="RNA-seq", ylab="Affymetrix array", main="PAX8 log2 expression")
```

## Create a meta-analysis dataset from curatedOvarianData

```{r createCODdata, cache=TRUE}
source(system.file("extdata", "patientselection.config",package="curatedOvarianData"))
rule.2 <- c("histological_type","^ser$")
rule.3 <- c("summarystage","^late$")
rule.4 <- c("summarygrade","^high$")
# add Surv objects as phenoData label "y" to the esets
add.surv.y <- function(X) Surv(X$days_to_death, X$vital_status=="deceased")
#if strict.checking is TRUE, patients missing any of the above
#metadata will be remove.  If FALSE, these patients are kept.
strict.checking <- TRUE
source(system.file("extdata", "createEsetList.R", package = "curatedOvarianData"))
esets <- esets[!grepl("TCGA.RNASeqV2_eset", names(esets))]
sum(sapply(esets, ncol))  #1427 samples
```


## Survival analysis

```{r, message=FALSE}
forestplot <- function(esets, y="y", probeset, formula=y~probeset,
                       mlab="Overall", rma.method="FE", at=NULL,xlab="Hazard Ratio",...) {
    require(metafor)
    esets <- esets[sapply(esets, function(x) probeset %in% featureNames(x))]
    coefs <- sapply(1:length(esets), function(i) {
        tmp <- as(phenoData(esets[[i]]), "data.frame")
        tmp$y <- esets[[i]][[y]]
        tmp$probeset <- exprs(esets[[i]])[probeset,]
        summary(coxph(formula,data=tmp))$coefficients[1,c(1,3)]
    })
    res.rma <- metafor::rma(yi = coefs[1,], sei = coefs[2,],
                            method=rma.method)
    if (is.null(at)) at <- log(c(0.25,1,4,20))
    forest.rma(res.rma, xlab=xlab, slab=gsub("_eset$","",names(esets)),
               atransf=exp, at=at, mlab=mlab,...)
    return(res.rma)
}
res <- forestplot(esets=esets,probeset="PAX8",at=log(c(0.5,1,2,4)))
```

## See if PAX8 is bimodal in more datasets

```{r, fig.height=12, fig.width=7}
par(mfrow=c(4, 3))
for (i in 1:length(esets)){
    if ("PAX8" %in% featureNames(esets[[i]]))
        hist(exprs(esets[[i]])["PAX8", ], main=names(esets)[[i]])
}
```

Is PAX8 expression related to primary site? 

```{r}
data(GSE9891_eset)
pData(GSE9891_eset)[1, ]
boxplot(exprs(GSE9891_eset)["PAX8", ] ~ GSE9891_eset$arrayedsite)
```

## Potential PAX8 targets

```{r}
library(pheatmap)
mat <- data.frame(t(exprs(TCGA_eset)[c("SNAI1", "TWIST1", "TIMP3", "E2F1", "PAX8"), ]))
pheatmap(t(mat), scale="row")
plot(PAX8 ~ TWIST1, dat=mat)
```

Significance tests for correlation

```{r}
cor.test(mat$PAX8, mat$TWIST1) #significant
cor.test(mat$PAX8, mat$SNAI1)  #significant only in GSE9891
cor.test(mat$PAX8, mat$TIMP3)  #not significant
cor.test(mat$PAX8, mat$E2F1)  #barely significant / not significant in GSE9891
```

## miRNA using array data from curatedOvarianData

```{r}
data(TCGA.mirna.8x15kv2_eset)
mir939 <- exprs(TCGA.mirna.8x15kv2_eset)["hsa-miR-939", ]
pax8 <- exprs(TCGA_eset)["PAX8", ]
pax8 <- pax8[names(pax8) %in% names(mir939)]
mir939 <- mir939[match(names(pax8), names(mir939))]
identical(names(pax8), names(mir939))
plot(log2(pax8+1) ~ log2(mir939+1))
cor.test(pax8, mir939, method="spearman")  #no correlation
```

## What is correlated with PAX8 and TWIST1?

```{r}
pax8.cor <- cor(exprs(TCGA_eset)["PAX8", ], t(exprs(TCGA_eset)))[1, ]
twist1.cor <- cor(exprs(TCGA_eset)["TWIST1", ], t(exprs(TCGA_eset)))[1, ]
hist(pax8.cor)
hist(twist1.cor)
```

Heatmap of genes correlated to PAX8 and TWIST1:

```{r, fig.width=9, fig.height=9}
twist1.cluster <- names(twist1.cor)[twist1.cor > 0.5]
pax8.cluster <- names(pax8.cor)[twist1.cor > 0.5]
pheatmap(exprs(TCGA_eset)[unique(c("PAX8", "TWIST1", twist1.cluster, pax8.cluster)), ],
         show_colnames=FALSE)
```
